---
title: "Initial Korean covid 19"
author: "mariahlynn"
date: "may 30,2024"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(digits=2)
require(tidyverse)
require(tigerstats)
require(lubridate)
kRdata <- read.csv(file = 'kRdata.csv')
kRdata <- kRdata %>%
mutate(across(where(is.character), as.factor))

```

# Introduction



### Data is from the Patient_Info dataset from South Korea 1 Jan to 1 May 2020
### This data was retrieved from a downloadable zip file from kaggle

Some of the data was uncoded for sex - sex must be just male or female
77 of the data entries were coded sex  as blank, so we omit these.
state of a patient must be died or alive - we adjust that as well   
  
# Your Question - Is dying independent of sex?


# Methods

result is a categorical variable, sex is categorical result depends on sex, we use helper file cat~cat




## Graphical results - use ggplot commands to do this  

```{r}
barchartGC(~result + Sex,data=kRdata)
barchartGC(~result + Sex,data=kRdata,  type="percent")
```
This barchart shows the frequency of women to men in comparison to being alive is greater , and being dead is lower than that of men. 

This barchart shows....
  





  
## Numerical results


```{r}
table1 <- xtabs (~result + sex, data=kRdata)
rowPerc(table1)
colPerc(table1)
```

this numerical summary shows that...
## Inferential results 




## Conclusions

