---
title: "Initial Korean covid 19"
author: "mariahlynn"
date: "may 30,2024"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(digits=2)
require(tidyverse)
require(tigerstats)
require(lubridate)
kRdata <- read.csv(file = 'kRdata.csv')
kRdata <- kRdata %>%
mutate(across(where(is.character), as.factor))

```

# Introduction



### Data is from the Patient_Info dataset from South   Korea 1 Jan to 1 May 2020  
### This data was retrieved from a downloadable zip file from kaggle  

Some of the data was uncoded for sex - sex must be just male or female  
77 of the data entries were coded sex  as blank, so we omit these.  
state of a patient must be died or alive - we adjust that as well    

# Your Question - Is dying independent of sex?

$H_0$: There is no difference in  .......    

$H_A$: There is a difference in.....    


# Methods 

result is a categorical variable, sex is categorical   
result depends on sex, we use helper file 
cat~cat  

# Results 


## Graphical results - use ggplot commands to do this

```{r}
barchartGC(~result + Sex,data=kRdata)
barchartGC(~result + Sex,data=kRdata, type="percent")

```
This barchart shows that...







## Numerical results 

```{r}
table1 <- xtabs(~result + Sex, data=kRdata)
rowPerc(table1)
colPerc(table1)
```
this numerical summary shows that...




## Inferential results


```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

The p-value for the Chi-squared test is 0.001, so we reject the null hypothesis.

The Fisher exact test shows that men are 2.3 tomes more likely to die than women in Korea. 


## Conclusions
The original question was, is there a relationship between sex and deaths in Korea due to covid. The p-value was 0.001, meaning that deaths caused by covid in Korea is not independent of gender . The Fisher test shows that men are 2.3 times more likely to die from covid compared to women in Korea. There is a relationship between gender or sex, and he amount of deaths in Korea during or due to Covid. 

